---
title: "Statistical inference with the GSS data"
output:
  word_document: default
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

## Introduction

Trust is essential to maintain financial system. Without it, we cannot
deposit money in a bank. Companies cannot borrow money from banks and
garner investment from people. The main reason of the Great Recession in
2008 was collective loss of trust in financial institutions.

In this brief paper, we will look at changes in people's trust in
financial institutions. We use the General Social Survey (GSS) to
investigate this. GSS is a collective data since 1972. It has monitored
social change in American society, including a question about the trust
in banks and other financial institutions.

From this data, we find more proportion of people who had a great deal
of confidence in financial institutions than that of people who had
hardly any confidence in them in the 1970s and most in the 1980s. In the
1990s, the proportions were fluctuating, then reversed in 2008, which
was exactly the year when the financial crisis occurred. We prove this
was not coincidental. In addition to this, we look at whether there was
any racial factor. We show that there was no racial bias affecting the
trust in financial institutions.

------------------------------------------------------------------------

## Part 1: Data

GSS data was collected by interviews whose median time was one and half
hour. Total rows are 57061, and 114 columns. The data includes 1972,
1973, 1974, 1975, 1976, 1977, 1978, 1980, 1982, 1983, 1984, 1985, 1986,
1987, 1988, 1989, 1990, 1991, 1993, 1994, 1996, 1998, 2000, 2002, 2004,
2006, 2008, 2010, 2012 data. Basically, the same questions were asked in
the survey, thus we look into annual changes in terms of each question.
In this paper, we will use tidyverse package and patchwork package in R.
The data was modified by Duke university from the original GSS to
exclude missing data and include factorized data.

```{r load-packages, message=FALSE, include=FALSE}
library(tidyverse)
library(patchwork)
load("gss.Rdata")
dim(gss)
unique(gss$year)
```

------------------------------------------------------------------------

## Part 2: Research question

1.  How has the trust in financial institutions changed in recent years?

2.  How did the financial crisis in 2008 affect people's confidence in
    financial institutions?

3.  Is there any differences in the confidence among race groups?

------------------------------------------------------------------------

## Part 3: Exploratory data analysis

First, we will illustrate annual changes in people's confidence in
financial institutions as a line graph. We drop years which do not
contain values (1972,73,74, and 85). We compare proportions of two
response, those who have a great deal of confidence in these
institutions and those who have hardly no confidence in them.

```{r echo=FALSE}

#replace NA with "No Answer"
levels <- levels(gss$confinan)
levels[length(levels) + 1] <- "No Answer"
gss$confinan <- factor(gss$confinan, levels = levels)
gss$confinan[is.na(gss$confinan)] <- "No Answer"

#make data frame including the annual changes
changes <- data.frame(addmargins(table(gss$year, gss$confinan)))
df_changes <- spread(data= changes,key =Var2 , value=Freq)
df_changes <- rename(df_changes, Year=Var1)
df_changes <- df_changes %>% mutate(pro_great=(`A Great Deal` /Sum), 
                                    pro_hardly=(`Hardly Any`/Sum))

#remove years that contain no data
df_changes <-df_changes[c(-1,-2,-3,-12,-30),] 

#graph 
g <-ggplot(data=df_changes) +
  geom_line(aes(Year,pro_great,group=1,col="A great deal"))
g+ geom_line(aes(Year,pro_hardly,group=1, col="Hardly any"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab("Year")+ylab("Confidence")+
  ggtitle("Changes in confidence on financial institutes")+
  geom_vline(xintercept = "2008",linetype=4)

```

From 1975 to 1989, the confidence in financial institutions were high,
even though we see gradual decrease throughout these years. From the
1990s, the figures had fluctuated until 2008. From 2008 to 2010, people
seemed to lose confidence in them considerably.

Then, we choose the data of 2008 and 2010, and compare them. Here are
the proportion tables, the upper is 2008 data and the lower is 2010
data.

```{r echo=FALSE}
#filter the data
gss_2010 <- gss %>% filter(year==2010)
gss_2008 <- gss %>% filter(year==2008) 

#cross tabulation
addmargins(prop.table(table("2008"=gss_2008$confinan)))*100
addmargins(prop.table(table("2010"=gss_2010$confinan)))*100
```

The proportion of high confidence("A Great Deal") dropped from 13.0% to
7.0 %, in contrast the proportion of no confidence("Hardly Any") almost
doubled( from 13.9% to 27.6%). The graphs clearly endorses these
figures.

```{r echo=FALSE, fig.height=5, fig.width=8}
g_2010 <- ggplot(data=gss_2010,aes(x=confinan))+
  geom_bar()+
  ggtitle("2010 Confidence in financial institution")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits = c(0, 900))

g_2008 <- ggplot(data=gss_2008,aes(x=confinan))+
  geom_bar()+
  ggtitle("2008 Confidence in financial institution")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits = c(0, 900))

g_2010 + g_2008

```

Did the financial crisis affect people's confidence in banks and other
institutions differently in term of races? We look into each response
further showing racial proportion. Here is the graph.

```{r echo=FALSE, fig.height=5, fig.width=8}
g_2010_fill <- ggplot(data=gss_2010,aes(x=confinan,fill=race))+
  geom_bar(position="fill")+
  ggtitle("2010 Confidence in financial institution")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g_2008_fill <- ggplot(data=gss_2008,aes(x=confinan,fill=race))+
  geom_bar(position="fill")+
  ggtitle("2008 Confidence in financial institution")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g_2010_fill + g_2008_fill

```

Show the racial proportion table of those who responded,"Hardly Any",
2008 and 2010 respectively;

```{r echo=FALSE}
prop.table(
  table("2008 'Hardly Any'"=
          (gss_2008 %>% filter(confinan=="Hardly Any"))$race))*100
```

```{r echo=FALSE}
prop.table(
  table("2010 'Hardly Any'"
        =(gss_2010 %>% filter(confinan=="Hardly Any"))$race))*100
```

The proportion of white increased (73.7% to 79.1%), however those of
black and other decreased(16.0% to13%, 10.3% to 7.4%, respectively).
Does that mean financial institutions were able to regain minority
groups' confidence?

Then look at the racial proportion table of those who responded,"A Great
Deal";

```{r echo=FALSE}
prop.table(
  table("2008 'A Great Deal'"
        =(gss_2008 %>% filter(confinan=="A Great Deal"))$race))*100
```

```{r echo=FALSE}
prop.table(
  table("2010 'A Great Deal'"
        =(gss_2010 %>% filter(confinan=="A Great Deal"))$race))*100
```

Just contrary to the previous response, white decreased (71.5% to
64.6%), black increased (16.0% to 22.9%), and other remained the same
(12.5%)

These two responses seem to indicate while the white lost confidence in
banks and other financial organizations, but the black regain confidence
somewhat by the financial crisis. As we know the crisis hit particularly
vulnerable people, the data seem counter-intuitive. We will also
investigate this statistically.

From these data exploration, we have two observations;

1.  People's confidence in financial institutions dropped from 2008 to
    2010
2.  Even though the confidence plummeted, there were racial differences.
    While whites seems to have been heavily hit, minority groups show a
    sign of improvement.

In the next segment, we will prove statistically whether the hypotheses
are true or not.

------------------------------------------------------------------------

## Part 4: Inference

Following studies, we set the significance level $\alpha=0.05$ .

```{r include=FALSE}
alpha =0.05

```

### **Significant changes between 2008 and 2010?**

#### **a) "A Great Deal" of confidence**

Here we will look into the response of those who have "A Great Deal" of
confidence in financial institutions. We look at the changes in the
confidence between 2008 and 2010. As we explored in the previous
segment, the descriptive data indicate the proportion of people who have
confidence in financial institution "a great deal" decreased. Do the
data provide convincing evidence to say this?

Before beginning statistical analysis, we check the conditions to apply
the normal models.

-   Independence:

The values are extracted from randomly selected data.

The sample size of 2008 ($n_{2008}=2023$) and 2010 ($n_{2010}=2044$) is
less than 10% of the population.

-   The success-failure condition:

$$
n_{2008} *p_{2008\,great}=263,\ n_{2008} *(1-p_{2008\,great})=1760\\
n_{2010} *p_{2010\,great}=144,\ n_{2010} *(1-p_{2010\,great})=1900
$$

All values are more than 10. The success- failure condition is met. We
assume these distributions are approximately normal.

```{r include=FALSE}
n_2008 <-nrow(gss_2008)
n_2010 <-nrow(gss_2010)
n_2008
n_2010

p_great_2008 = mean(gss_2008$confinan=="A Great Deal")
p_great_2010 = mean(gss_2010$confinan=="A Great Deal")
p_great_2008
p_great_2010


n_2008*p_great_2008
n_2008*(1-p_great_2008)
n_2010*p_great_2010
n_2010*(1-p_great_2010)

p_diff = p_great_2010 - p_great_2008
p_diff
```

**Confidence interval**

First, we calculate the 95% confidence interval for the point estimate,
the difference of the proportion between 2008 and 2010 values.

The sample proportions are,

$$
p_{great\,2008} = 0.1300049 \\
p_{great\,2010} =  0.0704501
$$

The point estimates, $$ { p_{diff}=-0.05955485 }$$.

The standard error,
$$ SE_{diff}= \sqrt{ \frac{p_{great\,2008}(1-p_{great\,2008})} {n_{2008}} + 
\frac{p_{great\,2010}(1-p_{great\,2010})} {n_{2010}} } = 0.009378031
$$

The 95% confidence interval is,
$$ p_{diff} \pm z_{0.975}*SE_{diff} = (-0.07793545, -0.04117424)$$

```{r include=FALSE}
se_ci <- sqrt(p_great_2008*(1-p_great_2008)/n_2008 +
                p_great_2010*(1-p_great_2010)/n_2010)

se_ci
ci_95 <-c(p_diff-qnorm(0.975)*se_ci, p_diff+qnorm(0.975)*se_ci)
ci_95
```

**Hypothesis testing**

Then, we conduct the hypothesis testing for the proportion of people who
have confidence in financial institution a great deal.

The null hypothesis is that there is no difference of the proportions
between 2008 and 2010.

The alternative hypothesis is that there is a difference between the
two.

$$
H_0: p_{2010\,great}=p_{2008\,great}\\H_A: p_{2010\,great}\neq p_{2008\,great}
$$

Pooled proportion,

$$
p_{pooled} =0.1000738
$$ Pooled standard error,

$$
SE_{pooled}= \sqrt{ \frac{p_{pooled}(1-p_{pooled})} {n_{2008}} + 
\frac{p_{pooled\,2010}(1-p_{pooled})} {n_{2010}} } = 0.009411574
$$

Given $H_0$ is true, p_value,

$$
p-value=2.486323e-10 \; <\alpha = 0.05
$$

P-value is far smaller than 0.05. We reject $H_0$. There is convincing
evidence of a difference between 2008 and 2010.

As we have seen the negative confidence interval, we understand that
people who have "A Great Deal" of confidence in financial institutions
**decreased** from 2008 to 2010.

```{r include=FALSE}
p_pool = (p_great_2008*n_2008 + p_great_2010*n_2010)/(n_2008+n_2010)
p_pool
se <- sqrt(p_pool*(1-p_pool)/n_2008 + 
             p_pool*(1-p_pool)/n_2010 )
se

z_sta <- p_diff/ se
z_sta
p_value <- pnorm(z_sta)*2
p_value
p_value < alpha

```

#### **b) "Hardly Any" confidence**

Next we will look into the response of those who have "Hardly any"
confidence in financial institutions. Likewise, We look at the changes
in the confidence between 2008 and 2010. As we explored in the previous
segment, the descriptive data indicate the proportion of people who
hardly have any confidence in financial institution increased.

We check the conditions to apply the normal models.

-   Independence:

The values are extracted from randomly selected data.

The sample size of 2008 ($n_{2008}=2023$) and 2010 ($n_{2010}=2044$) is
less than 10% of the population.

-   The success-failure condition:

$$
n_{2008} *p_{2008\,hardly}=281,\ n_{2008} *(1-p_{2008\,hardly}=1742\\
n_{2010} *p_{2010\,hardly}=564,\ n_{2010} *(1-p_{2010\,hardly})=1480
$$

All values are more than 10. The success- failure condition is met. We
assume these distributions are approximately normal.

```{r include=FALSE}

p_hardly_2008 = mean(gss_2008$confinan=="Hardly Any")
p_hardly_2010 = mean(gss_2010$confinan=="Hardly Any")
p_hardly_2008
p_hardly_2010


n_2008*p_hardly_2008
n_2008*(1-p_hardly_2008)
n_2010*p_hardly_2010
n_2010*(1-p_hardly_2010)

p_diff_h = p_hardly_2010 - p_hardly_2008
p_diff_h
```

**Confidence interval**

First, we calculate the 95% confidence interval for the point estimate,
the difference of the proportion between 2008 and 2010 values.

The sample proportions are,

$$
p_{hardly\,2008} =  0.1389026 \\
p_{hardly\,2010} =  0.2759295
$$

The point estimates, $$ { p_{diff}=0.1370269 }$$.

The standard error, $$ 
SE_{diff}= \sqrt{ \frac{p_{hardly\,2008}(1-p_{hardly\,2008})} {n_{2008}} + 
\frac{p_{hardly\,2010}(1-p_{hardly\,2010})} {n_{2010}} } = 0.01252478
$$

The 95% confidence interval is,
$$ p_{diff} \pm z_{0.975}*SE_{diff} = (0.1124788, 0.1615751)$$

```{r include=FALSE}
se_ci_h <- sqrt(p_hardly_2008*(1-p_hardly_2008)/n_2008 +
                p_hardly_2010*(1-p_hardly_2010)/n_2010)

se_ci_h
ci_95_h <-c(p_diff_h-qnorm(0.975)*se_ci, p_diff_h+qnorm(0.975)*se_ci)
ci_95_h
```

**Hypothesis testing**

Then, we conduct the hypothesis testing for the proportion of people who
hardly have any confidence in financial institutions.

The null hypothesis is that there is no difference of the proportions
between 2008 and 2010.

The alternative hypothesis is that there is a difference between the
two.

$$
H_0: p_{2010\,hardly}=p_{2008\,hardly}\\H_A: p_{2010\,hardly}\neq p_{2008\,hardly}
$$

Pooled proportion,

$$
p_{pooled} =0.2077699
$$ Pooled standard error, $$
SE_{pooled}= \sqrt{ \frac{p_{pooled}(1-p_{pooled})} {n_{2008}} + \frac{p_{pooled\,2010}(1-p_{pooled})} {n_{2010}} } = 0.01272377
$$

Given $H_0$ is true, p_value,

$$
p-value=4.802788e-27 \; <\alpha = 0.05
$$

P-value is far smaller than 0.05. We reject $H_0$. There is convincing
evidence of a difference between 2008 and 2010.

As we have seen the positive confidence interval, we understand that
people who have "Hardly Any" confidence in financial institutions
**increased** from 2008 to 2010.

```{r include=FALSE}
p_pool_h = (p_hardly_2008*n_2008 + p_hardly_2010*n_2010)/(n_2008+n_2010)
p_pool_h
se_h <- sqrt(p_pool_h*(1-p_pool_h)/n_2008 + 
           p_pool_h*(1-p_pool_h)/n_2010 )
se_h

z_sta_h <- p_diff_h/ se_h
z_sta_h
p_value_h <- pnorm(z_sta_h,lower.tail = F)*2
p_value_h
p_value_h < alpha
```

### **2 What races were affected by the financial crisis?**

From these studies above, we know that people had lost their confidence
in financial institutions from 2008 to 2010. In this segment, we look at
these differences in terms of racial proportion. In this segment, we
focus on people who have "Hardly Any" confidence in financial
institutions.

#### **a) "White" responded "Hardly Any"**

We check the conditions to apply the normal models.

-   Independence:

The values are extracted from randomly selected data.

The sample size of 2008 ($n_{2008\,hardly}=281$) and 2010
($n_{2010\,hardly}=564$) is less than 10% of the population.

-   The success-failure condition:

$$
n_{2008\,hardly} *p_{2008\,hardly\,white}=207,\ n_{2008\,hardly} *(1-p_{2008\,hardly\,white})=74\\
n_{2010\,hardly} *p_{2010\,hardly\,white}=446,\ n_{2010\,hardly} *(1-p_{2010\,hardly\,white})=118
$$

All values are more than 10. The success- failure condition is met. We
assume these distributions are approximately normal.

**Confidence interval**

First, we calculate the 95% confidence interval for the point estimate,
the difference of the proportion between 2008 and 2010 values.

The sample proportions are,

$$
p_{hardly\,white\,2008} =  0.7366548 \\
p_{hardly\,white\,2010} =  0.7907801
$$

The point estimates, $$ { p_{diff\,hardly\,white}=0.05412534 }$$.

The standard error, $$ 
SE_{diff\,hardly\,white}= \sqrt{ \frac{p_{hardly\,white\,2008}(1-p_{hardly\,white\,2008})} {n_{hardly\,2008}} + 
\frac{p_{hardly\,white\,2010}(1-p_{hardly\,white\,2010})} {n_{hardly\,2010}} } = 0.03136427
$$

The 95% confidence interval is,
$$ p_{diff\,hardly\,white} \pm z_{0.975}*SE_{diff\,hardly\,white} = (-0.007347504,  0.115598179)$$

This 95% confidence interval includes 0.

```{r include=FALSE}
#conditions

hardly_2008 <-gss_2008 %>% filter(confinan=="Hardly Any")
hardly_2010 <-gss_2010 %>% filter(confinan=="Hardly Any")

n_hardly_2008 <- nrow(hardly_2008)
n_hardly_2010 <- nrow(hardly_2010)
n_hardly_2008
n_hardly_2010

n_hardly_white_2008 <- sum(hardly_2008$race=="White")
n_hardly_white_2010 <- sum(hardly_2010$race=="White")
n_hardly_white_2008
n_hardly_white_2010


p_hardly_white_2008 = mean(hardly_2008$race=="White")
p_hardly_white_2010 = mean(hardly_2010$race=="White")
p_hardly_white_2008
p_hardly_white_2010

n_hardly_2008 * p_hardly_white_2008
n_hardly_2008 * (1- p_hardly_white_2008)

n_hardly_2010 * p_hardly_white_2010
n_hardly_2010 * (1- p_hardly_white_2010)



p_diff_hardly_white = p_hardly_white_2010 - p_hardly_white_2008
p_diff_hardly_white

se_hardly_white_ci <- sqrt((p_hardly_white_2008*
                              (1-p_hardly_white_2008)/n_hardly_2008+
                              (p_hardly_white_2010*
                                 (1-p_hardly_white_2010)/n_hardly_2010)))
se_hardly_white_ci

ci_95_w <- c(p_diff_hardly_white- qnorm(0.975)*se_hardly_white_ci,
             p_diff_hardly_white+ qnorm(0.975)*se_hardly_white_ci)
ci_95_w 

```

**Hypothesis testing**

Then, we conduct the hypothesis testing for the proportion of "white"
who hardly have any confidence in financial institutions.

The null hypothesis is that there is no difference of the proportions
between 2008 and 2010.

The alternative hypothesis is that there is a difference between the
two.

$$
H_0: p_{2010\,hardly\,white}=p_{2008\,hardly\,white}\\H_A: p_{2010\,hardly\,white}\neq p_{2008\,hardly\,white}
$$

Pooled proportion,

$$
p_{pooled\,hardly\,white} =0.7727811
$$ Pooled standard error, $$
SE_{pooled\,hardly\,white}= \sqrt{ \frac{p_{pooled\,hardly\,white}(1-p_{pooled\,hardly\,white})} {n_{2008\,hardly\,white}} + \frac{p_{pooled\,hardly\,white}(1-p_{pooled\,hardly\,white})} {n_{2010\,hardly\,white}} } = 0.03059751
$$ Given $H_0$ is true, p_value,

$$
p-value=0.07690295 \; >\alpha = 0.05
$$

P-value is larger than 0.05. We fail to reject $H_0$. There is not
convincing evidence of a difference of whites who responded that they
had hardly any confidence in financial institutions between 2008 and
2010.

As we have seen the confidence interval that include 0, we understand
that whites who have "Hardly Any" confidence in financial institutions
had **not changed** from 2008 to 2010.

```{r include=FALSE}
p_pool_hardly_white = (p_hardly_white_2008*n_hardly_2008 + 
                         p_hardly_white_2010*n_hardly_2010)/
                       (n_hardly_2008+n_hardly_2010)
p_pool_hardly_white
se_hardly_white <- sqrt(p_pool_hardly_white*(1-p_pool_hardly_white)/n_hardly_2008 +
                        p_pool_hardly_white*(1-p_pool_hardly_white)/n_hardly_2010 )

se_hardly_white
z_sta_hardly_white <- p_diff_hardly_white/ se_hardly_white
p_value_hardly_white <- pnorm(z_sta_hardly_white,lower.tail = F)*2
p_value_hardly_white

```

#### **b) "Black" responded "Hardly Any"**

We check the conditions to apply the normal models.

-   Independence:

The values are extracted from randomly selected data.

The sample size of 2008 ($n_{2008\,hardly}=281$) and 2010
($n_{2010\,hardly}=564$) is less than 10% of the population.

-   The success-failure condition:

$$
n_{2008\,hardly} *p_{2008\,hardly\,black}=45,\ n_{2008\,hardly} *(1-p_{2008\,hardly\,black})=236\\
n_{2010\,hardly} *p_{2010\,hardly\,black}=76,\ n_{2010\,hardly} *(1-p_{2010\,hardly\,black})=488
$$

All values are more than 10. The success- failure condition is met. We
assume these distributions are approximately normal.

**Confidence interval**

First, we calculate the 95% confidence interval for the point estimate,
the difference of the proportion between 2008 and 2010 values.

The sample proportions are,

$$
p_{hardly\,black\,2008} =  0.1601423 \\
p_{hardly\,black\,2010} =  0.1347518
$$

The point estimates, $$ { p_{diff\,hardly\,black}=-0.02539058 }$$.

The standard error, $$ 
SE_{diff\,hardly\,black}= \sqrt{ \frac{p_{hardly\,black\,2008}(1-p_{hardly\,black\,2008})} {n_{hardly\,2008}} + 
\frac{p_{hardly\,black\,2010}(1-p_{hardly\,black\,2010})} {n_{black\,2010}} } = 0.02617943
$$

The 95% confidence interval is,
$$ p_{diff\,hardly\,black} \pm z_{0.975}*SE_{diff\,hardly\,black} = (-0.07670132,  0.02592017)$$

This 95% confidence interval includes 0.

```{r include=FALSE}
#conditions

p_hardly_black_2008 = mean(hardly_2008$race=="Black")
p_hardly_black_2010 = mean(hardly_2010$race=="Black")
p_hardly_black_2008
p_hardly_black_2010

n_hardly_2008 * p_hardly_black_2008
n_hardly_2008 * (1- p_hardly_black_2008)

n_hardly_2010 * p_hardly_black_2010
n_hardly_2010 * (1- p_hardly_black_2010)



p_diff_hardly_black = p_hardly_black_2010 - p_hardly_black_2008
p_diff_hardly_black

se_hardly_black_ci <- sqrt((p_hardly_black_2008*
                              (1-p_hardly_black_2008)/n_hardly_2008+
                              (p_hardly_black_2010*
                                 (1-p_hardly_black_2010)/n_hardly_2010)))
se_hardly_black_ci

ci_95_b <- c(p_diff_hardly_black- qnorm(0.975)*se_hardly_black_ci,
             p_diff_hardly_black+ qnorm(0.975)*se_hardly_black_ci)
ci_95_b 

```

**Hypothesis testing**

Then, we conduct the hypothesis testing for the proportion of "black"
who hardly have any confidence in financial institutions.

The null hypothesis is that there is no difference of the proportions
between 2008 and 2010.

The alternative hypothesis is that there is a difference between the
two.

$$
H_0: p_{2010\,hardly\,black}=p_{2008\,hardly\,black}\\H_A: p_{2010\,hardly\,black}\neq p_{2008\,hardly\,black}
$$

Pooled proportion,

$$
p_{pooled\,hardly\,black} =0.1431953
$$ Pooled standard error,

$$
SE_{pooled\,hardly\,black}= \sqrt{ \frac{p_{pooled\,hardly\,black}(1-p_{pooled\,hardly\,black})} {n_{2008\,hardly\,black}} + \frac{p_{pooled\,hardly\,black}(1-p_{pooled\,hardly\,black})} {n_{2010\,hardly\,black}} } = 0.02557649
$$

Given $H_0$ is true, p_value,

$$
p-value= 0.320841 \; >\alpha = 0.05
$$

P-value is larger than 0.05. We fail to reject $H_0$. There is not
convincing evidence of a difference of blacks who responded that they
had hardly any confidence in financial institutions between 2008 and
2010.

As we have seen the confidence interval that include 0, we understand
that blacks who have "Hardly Any" confidence in financial institutions
had **not changed** from 2008 to 2010.

```{r include=FALSE}
p_diff_hardly_black
p_pool_hardly_black = (p_hardly_black_2008* n_hardly_2008 + 
                       p_hardly_black_2010* n_hardly_2010)/
                      (n_hardly_2008+n_hardly_2010)
p_pool_hardly_black

se_hardly_black <- sqrt(p_pool_hardly_black*(1-p_pool_hardly_black)/
                          n_hardly_2008 + 
                          p_pool_hardly_black*(1-p_pool_hardly_black)/
                          n_hardly_2010 )
se_hardly_black

z_sta_hardly_black <- p_diff_hardly_black/ se_hardly_black
z_sta_hardly_black
p_value_hardly_black <- pnorm(z_sta_hardly_black,lower.tail = T)*2
p_value_hardly_black
#we reject H_o
```

#### **c) "Other" responded "Hardly Any"**

We check the conditions to apply the normal models.

-   Independence:

The values are extracted from randomly selected data.

The sample size of 2008 ($n_{2008\,hardly}=281$) and 2010
($n_{2010\,hardly}=564$) is less than 10% of the population.

-   The success-failure condition:

$$
n_{2008\,hardly} *p_{2008\,hardly\,other}=29,\ n_{2008\,hardly} *(1-p_{2008\,hardly\,other})=252\\
n_{2010\,hardly} *p_{2010\,hardly\,other}=42,\ n_{2010\,hardly} *(1-p_{2010\,hardly\,other})=522
$$

All values are more than 10. The success- failure condition is met. We
assume these distributions are approximately normal.

**Confidence interval**

First, we calculate the 95% confidence interval for the point estimate,
the difference of the proportion between 2008 and 2010 values.

The sample proportions are,

$$
p_{hardly\,other\,2008} =  0.1032028 \\
p_{hardly\,other\,2010} =  0.07446809
$$

The point estimates, $$ { p_{diff\,hardly\,other}=-0.02539058 }$$.

The standard error, $$ 
SE_{diff\,hardly\,other}= \sqrt{ \frac{p_{hardly\,other\,2008}(1-p_{hardly\,other\,2008})} {n_{hardly\,2008}} + 
\frac{p_{hardly\,other\,2010}(1-p_{hardly\,other\,2010})} {n_{hardly\,2010}} } = 0.02125017
$$

The 95% confidence interval is,
$$ p_{diff\,hardly\,other} \pm z_{0.975}*SE_{diff\,hardly\,other} = (-0.07038433,  0.01291481)$$

This 95% confidence interval includes 0.

```{r include=FALSE}
#conditions

p_hardly_other_2008 = mean(hardly_2008$race=="Other")
p_hardly_other_2010 = mean(hardly_2010$race=="Other")
p_hardly_other_2008
p_hardly_other_2010

n_hardly_2008 * p_hardly_other_2008
n_hardly_2008 * (1- p_hardly_other_2008)

n_hardly_2010 * p_hardly_other_2010
n_hardly_2010 * (1- p_hardly_other_2010)



p_diff_hardly_other = p_hardly_other_2010 - p_hardly_other_2008
p_diff_hardly_black

se_hardly_other_ci <- sqrt((p_hardly_other_2008*
                              (1-p_hardly_other_2008)/n_hardly_2008+
                              (p_hardly_other_2010*
                                 (1-p_hardly_other_2010)/n_hardly_2010)))
se_hardly_other_ci

ci_95_o <- c(p_diff_hardly_other- qnorm(0.975)*se_hardly_other_ci,
             p_diff_hardly_other+ qnorm(0.975)*se_hardly_other_ci)
ci_95_o 

```

**Hypothesis testing**

Then, we conduct the hypothesis testing for the proportion of "other"
who hardly have any confidence in financial institutions.

The null hypothesis is that there is no difference of the proportions
between 2008 and 2010.

The alternative hypothesis is that there is a difference between the
two.

$$
H_0: p_{2010\,hardly\,other}=p_{2008\,hardly\,other}\\H_A: p_{2010\,hardly\,other}\neq p_{2008\,hardly\,other}
$$

Pooled proportion,

$$
p_{pooled\,hardly\,other} =0.08402367
$$ Pooled standard error, $$
SE_{pooled\,hardly\,other}= \sqrt{ \frac{p_{pooled\,hardly\,other}(1-p_{pooled\,hardly\,other})} {n_{2008\,hardly\,other}} + \frac{p_{pooled\,hardly\,other}(1-p_{pooled\,hardly\,other})} {n_{2010\,hardly\,other}} } = 0.02025716
$$

Given $H_0$ is true, p_value,

$$
p-value= 0.1560451 \; >\alpha = 0.05
$$

P-value is larger than 0.05. We fail to reject $H_0$. There is not
convincing evidence of a difference of others who responded that they
had hardly any confidence in financial institutions between 2008 and
2010.

As we have seen the confidence interval that include 0, we understand
that others who have "Hardly Any" confidence in financial institutions
had **not changed** from 2008 to 2010.

```{r include=FALSE}
p_diff_hardly_other
p_pool_hardly_other = (p_hardly_other_2008* n_hardly_2008 + 
                       p_hardly_other_2010* n_hardly_2010)/
                      (n_hardly_2008+n_hardly_2010)
p_pool_hardly_other

se_hardly_other <- sqrt(p_pool_hardly_other*(1-p_pool_hardly_other)/
                          n_hardly_2008 + 
                          p_pool_hardly_other*(1-p_pool_hardly_other)/
                          n_hardly_2010 )
se_hardly_other

z_sta_hardly_other <- p_diff_hardly_other/ se_hardly_other
z_sta_hardly_other
p_value_hardly_other <- pnorm(z_sta_hardly_other,lower.tail = T)*2
p_value_hardly_other
#we reject H_o
```

------------------------------------------------------------------------

## Conclusion

We illustrated three questions in the beginning of the paper;

1.  How has the trust in financial institutions changed in recent years?

2.  How did the financial crisis in 2008 affect people's confidence in
    financial institutions?

3.  Is there any differences in the confidence among race groups?

From 1975 to 1989, the confidence in financial institutions were high,
even though we see gradual decrease throughout these years. From the
1990s, the figures had fluctuated until 2008. From 2008 to 2010, people
seemed to lose confidence in them considerably.

From this, we focused on the changes between 2008 and 2010. As the
answer to the question 2 and 3, we found that people who had a great
deal of confidence in financial institutions plummeted significantly, at
the same time people who had hardly any confidence in them increased.
These are statistically proved. We assumed there might be racial
differences, but we could not find convincing evidence of racial
differences between 2008 and 2010. This implies that 2008 financial
crisis equally affected people notwithstanding their racial differences.

　
